---
title: "Practical 2 Answers"
subtitle: "Working with Data: Data Sources and Data Manipulation"
author: "Aysha Flueck"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: false
    link-external-newwindow: true
    link-external-icon: true
editor: source
---

```{=html}
<!--
Before you start:

- Make sure this file lives in the root of your project repository, alongside
  a `data/` folder containing:
    - reg-nurses-dataworld-1998-2020.csv
    - nurse-licenses-ncsbn-2026-01-05.csv
  (see the practical instructions for the download links).
- See the README.md in this repository for package installation instructions
  and Git/GitHub tips.
- Render this document with the "Render" button in RStudio, or by running
  quarto::quarto_render("practical2-answers.qmd") in the R console.
-->
```

## Getting started

```{r}
#| label: setup
#| message: false
#| warning: false

# Load the packages you need for this practical.
# See the README for installation instructions if any of these are missing.
library(readr)
library(dplyr)
library(ggplot2)
library(tibble)
library(tidyr)
```

## US State Nurse Data

### Question 1

Download the `.csv` file and use `readr::read_csv()` to read the dataset into R.

```{r}
#| label: read-data

nurses <- read_csv("data/reg-nurses-dataworld-1998-2020.csv")
```

### Question 2

Check the format of the data. Is it in a tidy format? Are there cleaning/tidying manipulations to do?

```{r}
#| label: inspect-data

glimpse(nurses)

# missingness
nurses |>
  summarise(across(everything(), ~mean(is.na(.)))) |> glimpse()

nurses |>
  filter(is.na(`Location Quotient`)) |> count(Year)

nurses |>
  filter(is.na(`Total Employed RN`)) |> select(State, Year)
```

**Is it tidy?** The data is mostly tidy. The rows and columns are organised in a logical way. However, there are some things to be improved:

- **Variable names** are messy.
- **Two columns are a different level of observation** (state/district level) and are the same for each entry within the 54 states/territories each year.
- **Location Quotient** is missing data before 2010, when a known change in reports occured. This missingness is structural.
- **Virgin Islands and Guam** have missing Total Employed RN entries in specific years. Pressumably structural.

```{r}
#| label: columns-rename

nurses <- nurses |>
  rename(
    total_employed        = `Total Employed RN`,
    employed_se_pct       = `Employed Standard Error (%)`,
    hourly_wage_avg       = `Hourly Wage Avg`,
    hourly_wage_median    = `Hourly Wage Median`,
    annual_salary_avg     = `Annual Salary Avg`,
    annual_salary_median  = `Annual Salary Median`,
    wage_se_pct           = `Wage/Salary standard error (%)`,
    hourly_p10            = `Hourly 10th Percentile`,
    hourly_p25            = `Hourly 25th Percentile`,
    hourly_p75            = `Hourly 75th Percentile`,
    hourly_p90            = `Hourly 90th Percentile`,
    annual_p10            = `Annual 10th Percentile`,
    annual_p25            = `Annual 25th Percentile`,
    annual_p75            = `Annual 75th Percentile`,
    annual_p90            = `Annual 90th Percentile`,
    location_quotient     = `Location Quotient`,
    national_total_employed          = `Total Employed (National)_Aggregate`,
    national_healthcare_employed     = `Total Employed (Healthcare, National)_Aggregate`,
    state_healthcare_employed        = `Total Employed (Healthcare, State)_Aggregate`,
    state_total_employed              = `Yearly Total Employed (State)_Aggregate`
  )

nurses
```

## Data Manipulation and Visualisation

### Question 3

Summarise the data available in the dataset. What years/states/variables are present?

```{r}
#| label: summarise-data

```

### Question 4

Using `dplyr::filter()` and `ggplot()`, recreate the plot from the 'Getting Started' section of the practical brief.

```{r}
#| label: recreate-plot

```

### Question 5

Create a new plot to visualise some other information (of your choice) from the dataset.

```{r}
#| label: own-plot

```

## Adding additional data

### Question 6

Read in the additional dataset from the National Council of State Boards of Nursing.

```{r}
#| label: read-additional-data
#| eval: false

nurse_licenses <- read_csv("data/nurse-licenses-ncsbn-2026-01-05.csv")
```

### Question 7

Summarise the data available in this new dataset. What years/states/variables are present?

```{r}
#| label: summarise-additional-data

```

### Question 8

What data is overlapping between this new dataset and the original one? Think about the temporal overlap and which types of nurses are present in each dataset.

### Question 9

If it is appropriate, create a combined dataset using both sources of information. If not appropriate, explain why.

```{r}
#| label: combine-data

```

### Question 10

Create a visualisation (of your choice) from the new or combined dataset.

```{r}
#| label: plot-new-or-combined

```

## Merging additional data

### Question 11

Filter the `reg-nurses-dataworld-1998-2020` data to only include 2017.

```{r}
#| label: filter-2017

```

### Question 12

Load the `us_rent_income` data located in the `tidyr` package.

```{r}
#| label: load-rent-income

# data(us_rent_income, package = "tidyr")
```

### Question 13

How would it be possible to merge the two datasets?

### Question 14

Merge the two datasets using `dplyr::full_join()`.

```{r}
#| label: merge-rent-nurses

```

### Question 15

Visualise the relationship between rent and nurse wages in 2017 using the combined data.

```{r}
#| label: plot-rent-vs-wages

```
